Verifying Spatial Properties of Array Computations

### 2017-04-17 oopsla17/algorithms.tex
Typo in nonOverlap
Add formatting break algorithm into two

\begin{enumerate}[leftmargin=1.5em]
\item (\textbf{Covering})
A covering of (possibly overlapping) the subscript space
is calculated as a value of $Region_N$. This is computed iteratively
by the following steps until a fixed-point is reached:

\begin{enumerate}[leftmargin=0.5em]
\item (\textbf{Coalesce intervals})
%
Index schemes which are \emph{contiguous} in one dimension are
coalesced.

\newcommand{\contig}[2]{\mathsf{contig}(#1, #2)}
\begin{defn}{[Contiguous]}
Two index schemes $S$ and $T$ are \emph{contiguous}
written $\contig{S}{T}$,
iff $\pi_i(S) = \pi_i(T)$
for all $1 \leq i \leq N$ apart from some dimension $k$ such that
$\pi_k(S) = [a, b]$ and $\pi_k(T) = [b+1, c]$.
\end{defn}
%
A coalesced set of index schemes is computed by:%
\begin{align*}
\mathsf{coalesce1}(S, \mathcal{S}) & =
 \{\,S \cup T \mid
T \in \mathcal{S} \, \wedge \,
\contig{S}{T} \} \\
\mathsf{coalesce}(\mathcal{S}) & =
\bigcup_{S \in \mathcal{S}} \mathsf{coalesce}(
\mathsf{coalesce1}(S, \mathcal{S})
\cup (\mathcal{S} - \{S\})) \\
\mathsf{coalesceTop}(\mathcal{S}) & =
 \mathsf{coalesce}(\mathcal{S})
\; \cup \;
\{ S \in \mathcal{S} \mid \neg \exists T \in \mathcal{S} . \; \contig{S}{T}\}
\end{align*}
\dnote{May need to prove that this is total}
%\begin{align*}
%\mathsf{coalesce}(\mathcal{S}) = \; &
%\{\,S \cup T \mid
%\forall S \in \mathcal{S}, \;
%\forall T \in \mathcal{S}. \;
%\contig{S}{T} \} \\
%\cup \; &
%\{ \,\;\;\; S \;\;\; \mid \forall S \in \mathcal{S},
%\; \not \exists T \in \mathcal{S} . \; \contig{S}{T}\}
%\end{align*}
%
That is, the new set of index schemes $\mathsf{coalesce}(\mathcal{S})$
comprises the union of the set of coalesced pairs of contiguous
schemes and the set of index schemes for which there is no other
contiguous scheme.
\item (\textbf{Remove redundant schemes})
Any indexing schemes which are subsumed by another are filtered out:
%
\begin{align*}
\textsf{nonOverlap}(\mathcal{S}) =
\{S \mid \forall S \in \mathcal{S}, \; \not \exists T \in (\mathcal{S} -
  \{S\}), \; \forall i . \; \pi_i(S) \subseteq \pi_i(T) \}
\end{align*}
%
\dnote{I'm starting to suspect that this step is actually redundant
  now- but I will need to do some math to check}
%\item \textbf{Covering}
\end{enumerate}
The least fixed-point of $\textsf{nonOverlap} \circ \textsf{coalesce}$
is computed, providing a minimal indexing scheme covering the space.
\dnote{Proof?}

For our example:
%
\begin{align*}
\textsf{coalesce}(\mathcal{S}_0) & =
\{[-1,0] \times [0,0],\,[0,0] \times [-1, 0],\,[0,0] \times [0,1],\,[0,1]
  \times [0,0]\} \\
%\textsf{nonOverlap}(\textsf{coalesce}(\mathcal{S}_0)) & =
%                                                        \textsf{coalesce}(\mathcal{S}_0)
%  \\
\textsf{coalesce}^2(\mathcal{S}_0) & =
\{[-1, 1] \times [0, 0],\,[0, 0] \times [-1, 1]\}
\end{align*}
%
The fixed-pointed is reached within two applications of
\textsf{coalesce} for our example.
\dnote{We probably need to say something about classifying index
schemes are \textit{Region$_N$}?}
\item (\textbf{Index schemes to syntax}) \dnote{Mistral todo}
\end{enumerate}

>>>>

We explain the algorithm in two parts. First, we show how \emph{adjacent} index
schemes are coalesced into smaller index schemes while remaining in union normal
form. Then, we explain how this smaller unioned indexing schemes are converted
to specifications.

\subsubsection{Covering}
A covering of (possibly overlapping) the subscript space
is calculated as a value of $Region_N$. This is computed iteratively
by the following steps until a fixed-point is reached:

\begin{description}[leftmargin=0.5em]

\item[Coalesce intervals]
%
Index schemes which are \emph{contiguous} in one dimension are
coalesced.

\newcommand{\contig}[2]{\mathsf{contig}(#1, #2)}
\begin{defn}{[Contiguous]}
Two index schemes $S$ and $T$ are \emph{contiguous}
written $\contig{S}{T}$,
iff $\pi_i(S) = \pi_i(T)$
for all $1 \leq i \leq N$ apart from some dimension $k$ such that
$\pi_k(S) = [a, b]$ and $\pi_k(T) = [b+1, c]$.
\end{defn}
%
A coalesced set of index schemes is computed by:%
\begin{align*}
\mathsf{coalesce1}(S, \mathcal{S}) & =
 \{\,S \cup T \mid
T \in \mathcal{S} \, \wedge \,
\contig{S}{T} \} \\
\mathsf{coalesce}(\mathcal{S}) & =
\bigcup_{S \in \mathcal{S}} \mathsf{coalesce}(
\mathsf{coalesce1}(S, \mathcal{S})
\cup (\mathcal{S} - \{S\})) \\
\mathsf{coalesceTop}(\mathcal{S}) & =
 \mathsf{coalesce}(\mathcal{S})
\; \cup \;
\{ S \in \mathcal{S} \mid \neg \exists T \in \mathcal{S} . \; \contig{S}{T}\}
\end{align*}
\dnote{May need to prove that this is total}
%\begin{align*}
%\mathsf{coalesce}(\mathcal{S}) = \; &
%\{\,S \cup T \mid
%\forall S \in \mathcal{S}, \;
%\forall T \in \mathcal{S}. \;
%\contig{S}{T} \} \\
%\cup \; &
%\{ \,\;\;\; S \;\;\; \mid \forall S \in \mathcal{S},
%\; \not \exists T \in \mathcal{S} . \; \contig{S}{T}\}
%\end{align*}
%
That is, the new set of index schemes $\mathsf{coalesce}(\mathcal{S})$
comprises the union of the set of coalesced pairs of contiguous
schemes and the set of index schemes for which there is no other
contiguous scheme.

\item[Remove redundant schemes]
Any indexing schemes which are subsumed by another are filtered out:
%
\begin{align*}
\textsf{nonOverlap}(\mathcal{S}) =
\{S \mid \forall S \in \mathcal{S}, \; \neg \exists T \in (\mathcal{S} -
  \{S\}), \; \forall i . \; \pi_i(S) \subseteq \pi_i(T) \}
\end{align*}
%
\dnote{I'm starting to suspect that this step is actually redundant
  now- but I will need to do some math to check}
%\item \textbf{Covering}
\end{description}

The least fixed-point of $\textsf{nonOverlap} \circ \textsf{coalesce}$
is computed, providing a minimal indexing scheme covering the space.
\dnote{Proof?}

For our example:
%
\begin{align*}
\textsf{coalesce}(\mathcal{S}_0) & =
\{[-1,0] \times [0,0],\,[0,0] \times [-1, 0],\,[0,0] \times [0,1],\,[0,1]
  \times [0,0]\} \\
%\textsf{nonOverlap}(\textsf{coalesce}(\mathcal{S}_0)) & =
%                                                        \textsf{coalesce}(\mathcal{S}_0)
%  \\
\textsf{coalesce}^2(\mathcal{S}_0) & =
\{[-1, 1] \times [0, 0],\,[0, 0] \times [-1, 1]\}
\end{align*}
%
The fixed-pointed is reached within two applications of
\textsf{coalesce} for our example.
\dnote{We probably need to say something about classifying index
schemes are \textit{Region$_N$}?}

\subsubsection{Index schemes to syntax}
